---
title: "STATISTICAL COMPUTING"
subtitle: "Getting to know RStudio"
author: "Austin Hart"
institute: "American University"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rladies-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  message=FALSE, warning=FALSE, eval = TRUE, echo = TRUE, 
  fig.align = 'center', dev='png', dpi = 300
)
knitr::opts_knit$set(root.dir = "C:/Users/ahart/Documents")
```

class: inverse, right, middle

# Understanding RStudio
Our analytic workhorse


``` {css, echo =F} 
    .pull-left {
      float: left;
      width: 35%;
    }

    .pull-right {
      float: right;
     width: 60%;
    }

    .pull-right ~ p {
      clear: both;
    }
```

---

# Materials and following along

- To particpate during/after session, download:
  - `womenparl.csv` dataset
  - `quickRscript.R` script
  
- How to manage the screens
  - Watch now; on your own later
  - Execute script as I go
  

---

# What is R/RStudio?

- **R**
  - Language/"environment" for statistical computing
  - Built specifically for data analysis, graphing
  - Easy extension via "packages"  
  
- **RStudio**
  - Friendly interface for R
  
- **Why bother?**
  - Powerful
  - Community-built
  - FREE! Free, free, free. 


---

# Think of it like a laptop

- **R**: computer out of the box  

- **RStudio**: your OS (Windows/MacOSX) for ease of use  

- **Packages**: the "Apps"/software you add on



---
class: left,middle

# Have a look around


---
class: inverse, right, middle

# How we interact with R

Basics of a session


---

# Typical analysis session in 600


1. Download data to project folder

2. Open RStudio; create new script
  - set directory
  - load packages
  
3. Write/execute code
  - import the data to RStudio
  - analyze, estimate, graph, etc 
  
4. Write up results in, e.g., Google Docs
  

---

# You NEED a project folder

> The working directory is a file path that sets the location for files you 
> load into or save out of R. If you try to load a data set, R assumes that 
> file is in your working directory. 


```{r}

# 1. Check my directory
  getwd()

# 2. Set the directory
  setwd("~/SIS600")
  
# 3. Is my data in there?
  list.files()
```


---

# An R Script file

.pull-left[

- Load packages  

- Set directory  

- Import data  

- Analysis...

]

.pull-right[

```{r scripting, eval = FALSE}
# PS 2, SIS600
# A. Hart
# ------------

# Activate packages
  library(tidyverse)

# Set directory
  setwd("~/SIS600")
  
# Load data
  load("DCPS testing.rdata")
  
# Q1: Mean of ProfLang
  summary(dcps$ProfLang)
```

]


---

# End of a session

- Save your script file!

- Typically no need to save the data

- We'll never need to save the "workspace"


---

# Coping with R

What to do when things go wrong?

- Check your script for errors/typos

- Check the R Guidebook!

- Search online (beware the rabbit hole)

- Ask questions effectively
  - Describe your issue and remedies attempted
  - Send screenshot/text of error message/etc


---
class: inverse, right, middle

# Quick example
Summary stats

---

# Women's representation in parliament

> `womenparl.csv`: Cross-national data on women's representation (percentage of seats held by women in lower house of national parliament).

- How many variables/obs in the data? 

- What percentage use a PR system?

- Describe data on `WomenParliament`
  - Present a graph 
  - Sentence or two with appropriate summary stats.  


---

# Extra credit!

> Is women's representation higher on average in PR systems than in non&ndash;PR systems? 

Extra credit point if you email me with:  

- Appropriate graph (1) of the group comparison (PR vs non&ndash;PR)  

- Short paragraph describing the difference with reference to appropriate summary statistics.  

- Replication code you used for the graph/calculations


2 points if you format the graph professionally!


---
class: inverse, middle

# Coming March 15!

- Workshop: racial resentment
  - Read Tesler chapters
  - Answer workshop prep questions  
  
- R skills builder quiz
  - Available on Blackboard
  - Attempt up to five (5) times


