---
title: "STATISTICAL COMPUTING"
subtitle: "Getting to know RStudio"
author: "Austin Hart"
institute: "American University"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rladies-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  message=FALSE, warning=FALSE, eval = TRUE, echo = TRUE, 
  fig.align = 'center', dev='svglite', dpi = 300
)
```


# Understanding RStudio
### Our analytic workhorse

- To participate, download:
  - `DCPS testing.rdata` data
  - `womenparl.Rdata` data
    
- For later: you need the R Survival Guide


``` {css, echo=F} 
.right-column{
  padding-top: 0;
}
```


---

# What is R/RStudio?

- **R**: Language/"environment" for stat computing
  - Built for data analysis, graphing
  - Easy extension via user-written "packages"  
  
- **RStudio**: Friendly interface for R
  
- **Why bother?**
  - Powerful
  - Community-built
  - FREE! Free, free, free. 


---

# Think of it like a laptop

- **R**: computer out of the box  

- **RStudio**: your OS (Windows/MacOSX) for ease of use  

- **Packages**: the "Apps"/software you add on



---
# Have a look around

- Fire up RStudio
  
- Install necessary packages (`tidyverse`, `stargazer`)
  
- Execute a line of code
  
- Close out


---
# Typical analysis session in 600

- Download data to project folder
  
- Open RStudio; create new script
  
- Write/execute code
  - load packages
  - set working directory
  - import the data to RStudio
  - analyze, estimate, graph, etc 
  
- Write up results
  
- Close the session
  

---
# You NEED a project folder

> The working directory is a file path that sets the location for files you 
> load into or save out of R. If you try to load a data set, R assumes that 
> file is in your working directory. 


.pull-left[
- I downloaded my data  
  
- Saved it in a special folder `Documents/SIS600`  
  
- Can I find it?
]

.pull-right[


```{r,eval=F}

# 1. Check my directory
  getwd()

# 2. Set my directory
  setwd("~/SIS600")
  
# 3. Is my data in there?
  list.files()
```
]

---

# An R Script file

.left-column[

- Load packages  

- Set directory  

- Import data  

- Analysis...

]

.right_column[

```{r scripting, eval = FALSE}
  # My first R script
  # A. Hart
  # .................
  
  # Setup
    library(tidyverse)
    setwd("~/SIS600")
    load("DCPS testing.rdata")
    
  # Q1: Describe ProfLang
    summary(dcps$ProfLang)
    hist(dcps$ProfLang)
```

]


---

# End of a session

- Save your script!

- Typically no need to save the data

- No need to save the "workspace"


---

# Coping with R

What to do when things go wrong?

- Check your script for errors/typos

- Check the R Guidebook!

- Ask questions effectively
  - Describe your issue and remedies attempted
  - Send screenshot/text of error message/etc
  
- Check online



